---
title: "Time Series Data Cleaning"
author: "Matthew Ross"
date: "4/20/2022"
output: html_document
---





# Time Series

Time series data, especially sensor water quality data is frequently gappy, prone
to sensor errors, and can be unintuitive to work with. Here we will
walk through some water quality data from the Kansas River in Topeka, Kansas.
This is an excellent site for exploring  water quality data because it is a
USGS 'Super gage" where they measure discharge, temp, do, chlorophyll a, salinity,
and turbidity. 

```{r}
library(tidyverse)
library(tsibble)
library(dataRetrieval)
library(mapview)
library(imputeTS)

```



# Download chlorophyll a data


```{r}

chl_dat <- readNWISuv('06888990',
                      parameterCd = c('32318'),
                      startDate = '2022-01-01',
                      endDate = '2022-03-01') %>%
  renameNWISColumns(p32318 = 'chl_a') %>% # A new magic function that gives things sensible names!
  #Ignore all data quality code columns 
  dplyr::select(-ends_with('_cd'))


```


## Plot chl_a data
```{r}
ggplot(chl_dat, aes(x=dateTime,y=chl_a_Inst)) + 
  geom_line() 

```

## Why that straight line? 

```{r}
ggplot(chl_dat, aes(x=dateTime,y=chl_a_Inst)) + 
  geom_line() + 
  geom_point(shape = 1)

```


# Making Implicit NAs explicit



## Convert to time-series tibble (tsibble)

[help here](https://tsibble.tidyverts.org/articles/implicit-na.html)


```{r}

chl_ts <- as_tsibble(chl_dat,
                     index = dateTime)
#Are there implicit gaps in the data? 
has_gaps(chl_ts)
```

### Where are the gaps in the data

```{r}
count_gaps(chl_ts)
```


### Fill the gaps with NAs

```{r}
chl_full <- chl_ts %>%
  fill_gaps()


ggplot(chl_full, aes(x = dateTime, y= chl_a_Inst)) + 
  geom_line()

#YAYYYYAYAYA
```


# Imputing over the gaps

```{r}


chl_gap_filled <- chl_full %>%
  mutate(impute_lin = na_interpolation(chl_a_Inst, 'linear'),
         impute_spline = na_interpolation(chl_a_Inst, 'spline'),
         impute_kalman = na_kalman(chl_a_Inst),
         impute_locf = na_locf(chl_a_Inst)) %>%
  pivot_longer(chl_a_Inst:impute_locf, names_to = 'imputation_method')



ggplot(chl_gap_filled, aes(x = dateTime, y=value, color = imputation_method,
                           size = imputation_method)) + 
         geom_line() +
  scale_size_manual(values = c(2,1.5,1,0.5,0.25)) + 
  ylim(0,50)


```



# Assignment!

## Repeat the above steps but use a different parameter (temp, turbidity, etc...)





# Downloading more complete datasets

Here we will be primarily exploring chlorophyl a data, which is a proxy for
algal biomass. We will also be downloading Discharge, temperature, and turbidity
data.



```{r}


top_dat <- readNWISuv('06888990',
                      parameterCd = c('00060','32318','00010',
                                      '00300','63680'),
                      startDate = '2022-01-01',
                      endDate = '2022-03-01') %>%
  renameNWISColumns(p32318 = 'chl_a') %>% # A new magic function that gives things sensible names!
  #Ignore all data quality code columns 
  dplyr::select(-ends_with('_cd'))


# Check out our data!
summary(top_dat)
```


## Reorganize and plot data


```{r}

top_long <- pivot_longer(top_dat,Wtemp_Inst:Turb_Inst,names_to = 'parameter')

ggplot(top_long, aes(x=dateTime,y=value)) + 
  geom_line() + 
  facet_wrap(~parameter, ncol=1, scales= 'free_y')


```


## Is 

```{r}

```

